### 2024-01-26 paper/paper.tex
20240126

\documentclass[ams]{U-AizuGT}
\usepackage{pifont}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{mathtools}
\usepackage{listings,jvlisting}

\lstset{
    frame=single,
    numbers=left,
    tabsize=2
}
\bibliographystyle{ieice}
\author{Ryuki Hiwada}
\studentid{s1280076}
\supervisor{Naohito Nakasato}

\title{Domain Specific Language for high performance computing}
\begin{document}
\maketitle
\section{Abstract}
\section{Introduction}
\subsection{Background}
In astrophysics and astronomy, to numerically calculate the dynamical 
evolution of N particles interacting gravitationally, N-body simulations 
are required. Figure 1 shows the equation for interparticle interactions 
in N-body simulations. If the equation is naively computed, the time
complexity of calculation of interparticle interactions is 
\begin{math}O(N^2) \end{math}, where 
N is the number of particles. Therefore, parallelization is required to
speed up numerical simulations. To write a parallelized code for a 
numerical simulation, a user needs to understand the architecture of 
computer systems in  detail. If a parallelized code is automatically
generated by only describing the formulas and data of the numerical
simulation, the above problems are solved. To realize the parallelization,
we will develop Domain Specific Language \lparen DSL \rparen, which is a 
programming language specialized to a domain, for example, SQL and HTML.
\subsection{parallelization}
SIMD is one of the categories in Flynn's taxonomy 1, related to computer arithmetic
 processing. Using SIMD, arithmetic operations can be applied simultaneously to
multiple data. In this paper, we compute and parallelize multiple particles. For example, 
as shown in Figure 1, the computation of acceleration for different particles is independent
of each other, allowing the process to be parallelized. As a result, for instance Assuming 
that the variables for particles are double-precision floating-point numbers, we can execute 
an operation on four elements simultaneously using the AVX2 instruction, it is possible to 
perform operations on four elements. Therefore, compared to non-parallelized code, there is
a potential to accelerate the computation by four times. We will describe the design of a 
DSL in the next section to generate such parallelized code from the formulation of particle-particle interactions.
\subsection{Overview of DSL}

We have created a Domain Specific Language \lparen DSL \rparen named \lparen DSL name \rparen. This language's 
processor can read code describing particle interaction formulas and generate code 
that accelerates these computations. The language allows for the definition of 
variables and the description of interaction formulas. In the variable definition 
section, it's possible to define variables such as the mass and position of 
particles, variables for storing results, and other necessary variables. For the formula
part, we have enabled calculations including basic arithmetic operations, sqrt, power 
functions, and computing the norm of vectors. From this description, a kernel function 
is generated, which can then be called and used. We have created a Domain Specific 
Language \lparen DSL \rparen named \lparen DSL name\rparen. This language's processor can read code describing particle 
interaction formulas and generate code that accelerates these computations. The language allows
for the definition of variables and the description of interaction formulas. In the variable
definition section, it's possible to define variables such as the mass and position of 
particles, variables for storing results, and other necessary variables. For the formula part, 
we have enabled calculations including basic arithmetic operations, sqrt, power functions, and 
computing the norm of vectors. From this description, a kernel function is generated, which can then be called and used.
\subsection{Aim of this paper}

\section{implementation}
We use Sympy, a library in Python for implementing  the implementation of our \lparen DSL\rparen .
SymPy supports various operations, and the formulas using these operations are internally treated as 
syntax trees. We leveraged this functionality to read formulas and generate code capable of parallel execution.
For parallelization, we used an instruction set named Advanced Vector Extensions 2 \lparen AVX2\rparen. AVX2 is one of the 
extended instruction sets implemented in Intel's CPUs. The next section will explain how parallelization is carried out using AVX2.

\subsection{AVX2 for parallelization}
To perform SIMD operations, just like with regular computations, data is first loaded from memory, computed based on that data,
and the results are stored. However, in the case of SIMD, computations are performed using SIMD-specific registers, so the 
methods of loading, computing, and storing differ. When loading, multiple elements that will undergo the same instruction are
stored in a SIMD-specific register. These are then used to perform computations simultaneously, and the results are stored for
each. For instance, consider a calculation like C = X + Y, where elements of double-precision floating arrays x and y are computed
. When using AVX2 for SIMD, since the registers are 256 bits, four elements from x and y are loaded each. These are then added 
simultaneously, and the results are stored in the array C. When loading into SIMD registers, if it's a 1-dimensional array, 
contiguous memory access is sufficient, but for data structures like Array of Structures \lparen AoS \rparen shown in Figure a., elements need
to be loaded individually. For example, when considering the 3-dimensional coordinates of particles as double-precision floats 
in pos\lbrack n\rbrack \lbrack 3\rbrack  \lparen where n is the total number of particles, and pos\lbrack i \rbrack \lbrack 0\rbrack , pos\lbrack i\rbrack \lbrack 1\rbrack , 
pos\lbrack i\rbrack \lbrack 2\rbrack  represent the x, y, z coordinates 
of the i-th particle, respectively\rparen, and calculating the difference in x coordinates between particles i and j, it becomes dx = 
pos\lbrack i\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack . However, for parallelization with SIMD, desiring dx0 = pos\lbrack i\rbrack \lbrack 0\rbrack  - pos\lbrack j \rbrack \lbrack 0\rbrack , dx1 = pos\lbrack i+1\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack 
, dx2 = pos\lbrack i+2\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack , dx3 = pos\lbrack i+3\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack , 
the data are packed as \lparen pos\lbrack i\rbrack \lbrack 0\rbrack , pos\lbrack i+1\rbrack \lbrack 0\rbrack , pos\lbrack i+2\rbrack \lbrack  0\rbrack , pos\lbrack  i+3\rbrack \lbrack  0\rbrack \rparen,
\lparen pos\lbrack  j\rbrack \lbrack  0\rbrack , pos\lbrack  j\rbrack \lbrack  0\rbrack , pos\lbrack  j\rbrack \lbrack  0\rbrack , pos\lbrack  j\rbrack \lbrack  0\rparen, \lparen dx0, dx1, dx2, dx3\rparen and subtraction is performed simultaneously. However, since data needs
to be loaded onto SIMD registers all at once, the gather instruction is used to load data for particle i. The gather instruction is used
to read non-contiguous data elements from memory addresses by specifying addresses. An explanation of the gather instruction is shown in 
Figure x. Data loaded in this way are computed simultaneously using SIMD instructions. When storing computation results, addresses may 
not be contiguous. In such cases, there is an instruction called scatter, which is the counterpart to gather, but it is not supported 
in AVX2, so data are stored using pointers individually. In this way, parallelization is achieved with SIMD. Moreover, these instructions
can be explicitly handled in high-level programming languages like C and C++ using intrinsic functions. To handle intrinsic functions 
in C and C++, include the file immintrin.h, which is available as a standard in the language. An example of C++ code that calculates the difference in x coordinates, simd\_tmp.cpp, is shown. { DSL name}  generates C++ code capable of executing SIMD instructions like simd\_tmp.cpp
from the described formulas. The next section will explain the design of { DSL name}.

\subsection{design of name of DSL}

\begin{lstlisting}[language=c++]
#include <iostream>

int main(){
    std::cout << "Hello, world!" << std::endl;
    return 0;
}
\end{lstlisting}
\subsection{implementation of (name of DSL)}

\subsection{Use Sympy for DSL development}

\section{Conclusion}
\section{Acknowledgement}
\bibliography{biblist}
\end{document}